---
title: "README.Rmd"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "img/README_")
library(printr)
```

# Ça va, CAVA? Dictionary Coherence, Augmentation, (Validation and Analysis)

CAVA is an R package to assit in working with dictionary (keywords/lexical text analysis) in a valid way.
It allows you to use an embeddings model to do dictionary expansion/augmentation, check its coherence, (and at some future date) validation and analysis.

For a longer description, see our [ICA Tool Demo abstract](CAVA_ICA_abstract.pdf).

## Installing and obtaining an embeddings model

You can install CAVA from github:

```{r, eval=F}
remotes::install_github("vanatteveldt/CAVA")
```

Before starting, you need an embeddings model. Currently, we only support Fasttext .bin models. 
The code below downloads an English fasttext model:

```{r}
if (!file.exists("cc.en.300.bin")) {
  url = "https://dl.fbaipublicfiles.com/fasttext/vectors-crawl/cc.en.300.bin.gz"
  options(timeout=4300)  # assuming 1Mb/s
  download.file(paste0(url), destfile = "cc.en.300.bin.gz")
  R.utils::gunzip("cc.en.300.bin.gz") # Install R.utils if needed
}
```

# Using CAVA

The main functions exposed to cava are shown below. 
For a more elaborate example, please see the [example usage file](usage.md).

Loading the FastText mnodel, using the state of the union speeches as target corpus:

```{r}
library(CAVA)
corpus = quanteda::corpus(sotu::sotu_text, docvars = sotu::sotu_meta)
vectors = load_fasttext("cc.en.300.bin", corpus)
```

## Augmentation

Expanding a dictionary using wildcard and similarity:

```{r}
dictionary = c("fin*", "eco*")
dictionary = expand_wildcards(dictionary, vectors)
candidates = similar_words(dictionary, vectors)
dictionary = c(dictionary, candidates$word[candidates$similarity>.4])
head(candidates)
```

Expanding a dictionary using antonyms:

```{r}
positive = c("good", "nice", "best", "happy")
negative = c("evil", "nasty", "worst", "bad", "unhappy")
candidates = similar_words(positive, vectors, antonyms = negative)
head(candidates)
```

## Coherence

Computing and plotting pairwise similarities:

```{r coherence}
similarities = pairwise_similarities(dictionary, vectors)
similarities |> similarity_graph(max_edges=100) |> plot()
```

Computing similarity to dictionary centroid (sorted with most distances words on top):

```{r}
similarity_to_centroid(dictionary, vectors) |> head()
```

